---
title: "607 Week3 ALiu-Ferrara"
author: "Ann Liu-Ferrara"
date: "February 18, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
```

## 3. Copy the introductory example. The vector name stores the extracted names.
R> name
[1] "Moe Szyslak" "Burns, C. Montgomery" "Rev. Timothy Lovejoy"
[4] "Ned Flanders" "Simpson, Homer" "Dr. Julius Hibbert"

(a) Use the tools of this chapter to rearrange the vector so that all elements conform to
the standard first_name last_name.
```{r echo=TRUE}
name <- c("Moe Szyslak", "Burns, C. Montgomery", "Rev. Timothy Lovejoy",
          "Ned Flanders", "Simpson, Homer", "Dr. Julius Hibbert")

 first <- unlist(str_extract_all(name, '\\w+ |\\, \\w+$|\\. \\w+$'))
 first <- unlist(str_extract_all(first, '\\w+'))
 last <- unlist(str_extract_all(name, '\\b \\w+|\\w+,'))
 last <- unlist(str_extract_all(last, '\\w+'))
 full_name <- paste0(first, ' ', last)
 full_name

```

(b) Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.).
```{r echo = TRUE}
title <- str_detect(name, 'Rev.|Dr.')
```

(c) Construct a logical vector indicating whether a character has a second name.
```{r echo = TRUE}
  second_name <- str_detect(name, '\\ .\\. ')
```

## 4. Describe the types of strings that conform to the following regular expressions and
construct an example that is matched by the regular expression.

(a) [0-9]+\\$

  Digits ending with dollar sign
```{r echo = TRUE}
unlist(str_extract_all('212$13 33$', '[0-9]+\\$'))
```
(b) \\b[a-z]{1,4}\\b

  Any lower case words with a word edge and repeated at least 1 and up to 4 times.
```{r echo = TRUE}
unlist(str_extract_all('a wordExample4 2 word', '\\b[a-z]{1,4}\\b'))
```
(c) .*?\\.txt$

  Either none or any characters ending with .txt
```{r echo = TRUE}
unlist(str_extract_all('a wordExample4 2 word.txt', '.*?\\.txt$'))
unlist(str_extract_all('.txt', '.*?\\.txt$'))

```
(d) \\d{2}/\\d{2}/\\d{4}

  Date format with '/', dd/mm/year
```{r echo = TRUE}
unlist(str_extract_all('23/02/1902 23/12/1998', '\\d{2}/\\d{2}/\\d{4}'))

```
(e) <(.+?)>.+?</\\1>

  Search for html or xml tag pattern begin with <tag> and end with </tag>
```{r echo = TRUE}
unlist(str_extract_all('<html>content</html>, <tag>text</tag>', '<(.+?)>.+?</\\1>'))

```

Reference: 
https://github.com/wwells/CUNY_DATA_607/blob/master/Week3/Regex_Week3.Rmd
