---
title: "607 Week3 ALiu-Ferrara"
author: "Ann Liu-Ferrara"
date: "February 18, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
```

## 3. Copy the introductory example. The vector name stores the extracted names.
R> name
[1] "Moe Szyslak" "Burns, C. Montgomery" "Rev. Timothy Lovejoy"
[4] "Ned Flanders" "Simpson, Homer" "Dr. Julius Hibbert"
```{r echo=FALSE}
name <- c("Moe Szyslak", "Burns, C. Montgomery", "Rev. Timothy Lovejoy",
          "Ned Flanders", "Simpson, Homer", "Dr. Julius Hibbert")
name
```

(a) Use the tools of this chapter to rearrange the vector so that all elements conform to
the standard first_name last_name.
```{r echo=FALSE}

```

(b) Construct a logical vector indicating whether a character has a title (i.e., Rev. and
Dr.).


(c) Construct a logical vector indicating whether a character has a second name.


33 4. Describe the types of strings that conform to the following regular expressions and
construct an example that is matched by the regular expression.

(a) [0-9]+\\$
(b) \\b[a-z]{1,4}\\b
(c) .*?\\.txt$
(d) \\d{2}/\\d{2}/\\d{4}
(e) <(.+?)>.+?</\\1>

```{r cars}

```
