---
title: "607_ALiu-Ferrara_Week2"
author: "Ann Liu-Ferrara"
date: "February 12, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment – SQL and R

Choose six recent popular movies. Ask at least five people that you know (friends, family, classmates, imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5. Take the  
results (observations) and store them in a SQL database. Load the information into an R dataframe. Your deliverables should include your SQL scripts and your R Markdown code, posted to GitHub.

```{r echo = TRUE}

#if(!require("RMySQL")) install.packages("RMySQL")
#if(!require("ggplot2")) install.packages("ggplot2")

library(RMySQL)
library(ggplot2)

mydb = dbConnect(MySQL(), user='xxxxx', password='xxxxx', dbname='movies', host='localhost')
dbListTables(mydb)
rs <- dbSendQuery(mydb, 'select * from movie_reviews')
movieRating <- fetch(rs, n = -1)

# The movie rating by reviewer
qplot(rating, movie, data = movieRating,  xlab = "Rating", ylab = "Movie",
      main = "Individual Movie Rating by Reviewer") +
  facet_wrap(~reviewer) 

# Movie rating from highest to lowest
ggplot(movieRating, aes(x = reorder(movie, rating), y = rating, fill = movie),
      main = 'Histogram for Rating', xlab = 'Rating', fill = I("lightblue"), col = I("grey")) +
  geom_bar(stat = "identity") + 
  ggtitle("Movie Ratings") +
  labs(x = "Movie") +
  coord_flip()

# Movie Rating by reviewer
ggplot(movieRating, aes(x = reorder(reviewer, -rating), y = rating, fill = reviewer)) + 
  geom_bar(stat = "identity") + 
  ggtitle("Total Movie Rating by Reviewer") + 
  labs(x = "Reviewer")

dbListConnections(dbDriver(drv = "MySQL"))
```
