---
title: "Assignment - Tidying and Transforming Data - Week 5"
author: "Ann Liu-Ferrara"
date: "March 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
The chart above describes arrival delays for two airlines across five destinations. Your task is to:
(1) Create a .CSV file (or optionally, a MySQL database!) that includes all of the information above. You're encouraged to use a "wide" structure similar to how the information appears above, so that you can practice tidying and transformations as described below.

(2) Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data.

```{r echo = TRUE}
library(dplyr)
library(tidyr)
library(ggplot2)

# read .csv file
(flight_df<- tbl_df(read.csv(file="fly.csv",head=T, sep=",")))

# tidy data by putting city names as values in column 'city', and set up 'delayed' and 'on time' as variable, and spread into columns. 
g <- flight_df %>%  
  gather(City, flights, 3:ncol(flight_df), -Status) %>%   
      spread(Status, flights)

# rename the last 2 variables
names(g)[3:4] <- c('Delayed', 'OnTime')
```

(3) Perform analysis to compare the arrival delays for the two airlines.

```{r echo=TRUE}

# compare delays using the mean, sum, min, and max data from both airlines.
g %>% 
  group_by(Company) %>%
      summarise_each(funs(mean, sum, min, max), Delayed) 
   
      
# visualize the delayed flights by city by company
ggplot(data = g) +
  geom_point(aes(x = City, y = Delayed), color = 'blue', size = 3) +
  facet_grid(. ~ Company, scales="free") +
  coord_flip()

```

Conclusion:

Overall comparison of the two airlines, ALASKA has less delays comparing to AM WEST with lower mean, totall delays, and the minimum and maximum delay records by cities.