---
title: "Week 13 assignment - NoSQL migration"
author: "Ann Liu-Ferrara"
date: "April 28, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# In this assignment, I will take flights data from a csv file and migrate it to MongoDB database, and describe the advantages and disadvantages of storing the data in a relational database vs. NoSQL database.

```{r echo= TRUE}
library(RMongo)
library(rjson)
# connect db
mongo <- mongoDbConnect('test', '127.0.0.1', 27017)

# read flights
flights <- read.csv("flights.csv", stringsAsFactors = FALSE)
L=lapply(split(flights,rownames(flights)),as.list)
names(L)=NULL

# remove all rows
# output <- dbRemoveQuery(mongo, 'flights', '{}')

# insert flights
for (i in 1:NROW(L)) {
dataJSON = toJSON(L[[i]])
output <- dbInsertDocument(mongo, "flights", dataJSON)
}

#output <- dbGetQuery(mongo, 'flights', '{"depart": "Seattle"}')
# view all records in table flights
output <- dbGetQuery(mongo, 'flights', '{}')
print(output)

# disconnect mongoDB
dbDisconnect(mongo)
flights <- flights[with(flights, order(arrive, depart, flighttime)), ]
output <- output[, -2]
output <- output[with(output, order(arrive, depart, flighttime)), ]
output <- output[, c(1, 3, 2)]
flights <- flights[, c(2, 1, 3)]

# compare 2 databases
all.equal(flights, output,check.attributes = FALSE)


```


# Advantages of NoSQL vs. relational database:

1, flexibility, easy to manage than relational database; 

2, mostly open source and low-cost; 

3, scalability in supporting Mapreduce; 

4, less complexity, no relational database set up; and

5, grow rapidly.


Disadvantages of NoSQL vs. relational database, such as MySQL:

1, Not as maturing as MySQL due to short history;

2, lack of reporting tools;

3, lack of standardization in querying language and compatibility; and

4, the community is not well defined.

Conclusion:

NoSQL data is the product of rapidly developed social media, big data, and web technologies, there is a need for more flexible structure and easier operation. It might not replace relational database, but grow as an important supplement to meet the demand in data storage.


References:

http://stackoverflow.com/questions/19564321/how-to-send-multiple-documents-using-rmongo

http://www.monitis.com/blog/cc-in-review-the-key-differences-between-sql-and-nosql-dbs/