---
title: "IS 605 FUNDAMENTALS OF COMPUTATIONAL MATHEMATICS - SPRING 2017, Assignment Week 2"
author: "Ann Liu-Ferrara"
date: "February 9, 2017"
output: 
  html_document:
    mathjax: "http://cdn.mathjax.org/mathjax/latest/MathJax.js?config=TeX-AMS-MML_HTMLorMML"
runtime: shiny
---

$\mathbf{X}^\top$ is the matrix transpose of $\mathbf{X}$.

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```
## Problem set 1
(1) Show that AT $A^TA$ $\neq$ $AA^T$ in general. (Proof and demonstration.)

        Proof:as AB $\neq$ BA, let B = $A^T$, then $A^TA$ $\neq$ $AA^T$

        Demonstration:

\[
\begin{pmatrix}
  \begin{bmatrix}
    1       & 2 & 3  \\
    4       & 5 & 6  \\
    7       & 8 & 9 
  \end{bmatrix}
\end{pmatrix}^T
\begin{bmatrix}
    1       & 2 & 3  \\
    4       & 5 & 6  \\
    7       & 8 & 9 
\end{bmatrix}
\stackrel{?}{=}
\begin{bmatrix}
    1       & 2 & 3  \\
    4       & 5 & 6  \\
    7       & 8 & 9 
\end{bmatrix}
\begin{pmatrix}
  \begin{bmatrix}
    1       & 2 & 3  \\
    4       & 5 & 6  \\
    7       & 8 & 9 
  \end{bmatrix}
\end{pmatrix}^T
\]


\[
\begin{bmatrix}
    66       & 78 & 90  \\
    78       & 93 & 108  \\
    90       & 108 & 126 
\end{bmatrix}
\neq
\begin{bmatrix}
    14       & 32 & 50  \\
    32       & 77 & 122  \\
    50       & 122 & 194
\end{bmatrix}
\]


(2) For a special type of square matrix A, we get $A^TA$ = $AA^T$. Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix). Please typeset your response using LaTeX mode in RStudio. 

        Identity and symmetric matrices equal to their inverses, so have the property AB = BA:

\[
\begin{pmatrix}
  \begin{bmatrix}
    1       & 0 & 0  \\
    0       & 1 & 0  \\
    0       & 0 & 1 
  \end{bmatrix}
\end{pmatrix}^T
\begin{bmatrix}
    1       & 0 & 0  \\
    0       & 1 & 0  \\
    0       & 0 & 1 
\end{bmatrix}
\stackrel{?}{=}
\begin{bmatrix}
    1       & 0 & 0  \\
    0       & 1 & 0  \\
    0       & 0 & 1 
\end{bmatrix}
\begin{pmatrix}
  \begin{bmatrix}
    1       & 0 & 0  \\
    0       & 1 & 0  \\
    0       & 0 & 1 
  \end{bmatrix}
\end{pmatrix}^T
\]


\[
\begin{bmatrix}
    1       & 0 & 0  \\
    0       & 1 & 0  \\
    0       & 0 & 1 
\end{bmatrix}
=
\begin{bmatrix}
    1       & 0 & 0  \\
    0       & 1 & 0  \\
    0       & 0 & 1 
\end{bmatrix}
\]

## Problem set 2
Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being factorized. Radars that track flights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your flight using radars. Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer. 

```{r echo=TRUE}
DecomposeLU <- function (A) {
  if(dim(A)[1] != dim(A)[2]) 
    return(NULL)

  U <- A
  L <- diag(dim(A)[1])
          
  # create E, and update E, U and L 
  for (i in 2:dim(A)[1]) {
    for (j in 1:(i - 1)) {
      E <- diag(dim(A)[1])
      E[i, j] <- -(U[i, j]/U[j, j])
      U <- E %*% U
      L <- L %*% solve(E)
    }
  }
  
  return(list(L, U))
}

```

Reference:https://github.com/wwells/CUNY_DATA_605/blob/master/Week02/WWells_Assign2.Rmd

