---
title: "IS 605 FUNDAMENTALS OF COMPUTATIONAL MATHEMATICS - SPRING 2017, Assignment Week 3"
author: "Ann Liu-Ferrara"
date: "February 21, 2017"
output: 
  html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```
## Problem set 1

(1) What is the rank of the matrix A?

  The rank of the matrix A is 4. It is a linearly independent matrix.

```{r echo=TRUE}
library(Matrix)
A <- matrix(c(1, -1, 0, 5, 2, 0, 1, 4, 3, 1, -2, -1, 4, 3, 1, -3), nrow = 4)
A
rankMatrix(A)
```
(2) Given an m x n matrix where m > n, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero?

  The maximum rank is the minimum of m and n, since m > n, the maximum rank is n. The minimum rank for a non-zero matrix is 1.

(3) What is the rank of matrix B?

  The rank of matrix B is 1, as matrix B is not a linearly independent matrix.
```{r echo=TRUE}
B <- matrix(c(1, 3, 2, 2, 6, 4, 1, 3, 2), nrow = 3)
B
rankMatrix(B)

```
## Problem set 2

Compute the eigenvalues and eigenvectors of the matrix A. You'll need to show your work. You'll need to write out the characteristic polynomial and show your solution.
```{r echo=TRUE}
A <- matrix(c(1, 0, 0, 2, 4, 0, 3, 5, 6), nrow = 3)
A
```

${det}\left(A-\lambda \,I\right)\,=\,0?$

The characteristic equation of the matrix

$\left(\begin{array}{ccc}1&2&3\\0&4&5\\0&0&6\end{array}\right)$ 

is

$\text{det}\left(\begin{array}{ccc}1-\lambda &2&3\\0&4-\lambda &5\\0&0&6-\lambda \end{array}\right)=0$

Expanding the determinant,

$\left(1-\lambda\right) [\left(4-\lambda \right)\left(6-\lambda \right)-5\times 0] + 2[0 \times \left(6-\lambda \right)-5\times 0] + 3[0\times0 - 0\left(4-\lambda \right)]=0$

$\lambda_1=6,\lambda_2=4,\lambda_3=1$

Compute the eigenvectors:

If $\lambda_1=6$ and set X = 1, then solve the equation:

$\left(\begin{array}{ccc}-5&2&3\\0&-2&5\\0&0&0\end{array}\right)\left(\begin{array}{c}X\\Y\\Z\end{array}\right)=\left(\begin{array}{c}0\\0\\0\end{array}\right).$

-5 + 2Y + 3Z = 0

-2Y + 5Z = 0

X = 1, Y = $\frac{25}{16}$, Z = $\frac{5}{8}$

Scale up, the eigenvector is (16, 25, 10)

Same calculations for $\lambda_2=4$ and $\lambda_3=1$, the eigenvectors are (2, 3, 0) and (1, 0, 0).

## Testing:
```{r echo=TRUE}
ev <- eigen(A)
# extract components
(values <- ev$values)

(vectors <- ev$vectors)
```

Reference: 

http://wwwf.imperial.ac.uk/metric/metric_public/matrices/eigenvalues_and_eigenvectors/eigenvalues2.html

https://cran.r-project.org/web/packages/matlib/vignettes/eigen-ex1.html


