---
title: "ASSIGNMENT 12"
author: "Ann Liu-Ferrara"
date: "April 19, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## BIAS VARIANCE TRADEOFF IN R

Using the stats and boot libraries in R perform a cross-validation experiment to observe the bias variance tradeoff. You'll use the auto data set from previous assignments. This dataset has 392 observations across 5 variables. We want to fit a polynomial model of various degrees using the glm function in R and then measure the cross validation error using cv.glm function.
Fit various polynomial models to compute mpg as a function of the other four variables acceleration, weight, horsepower, and displacement using glm function. For example: 

glm.fit=glm(mpg~poly(disp+hp+wt+acc,2), data=auto)
cv.err5=cv.glm(auto,glm.fit,K=5)$delta[1]

will fit a 2nd degree polynomial function between mpg and the remaining 4 variables and perform 5 iterations of cross-validations. This result will be stored in a cv.err5 array. cv.glm returns the estimated cross validation error and its adjusted value in a variable
called delta. Please see the help on cv.glm to see more information.
Once you have fit the various polynomials from degree 1 to 8, you can plot the crossvalidation error function as

degree=1:8
plot(degree, cv.err5, type='b')

For you assignment, please create an R-markdown document where you load the auto data set, perform the polynomial fit and then plot the resulting 5 fold cross validation curve.

Your output should show the characteristic U-shape illustrating the tradeoff between bias and variance.

```{r echo = TRUE}
library(stats)
library(boot)

# read in data
auto <- read.table('auto-mpg.data', quote='\'', comment.char='')
names(auto) <- c('disp', 'hp', 'wt', 'acc', 'mpg')
head(auto)

degree=1:8
cv.err5 <- NULL
for(i in degree) {
glm.fit <- glm(mpg~poly(disp+hp+wt+acc,i), data=auto)
cv.err5 <- c(cv.err5, cv.glm(auto,glm.fit,K=5)$delta[1])
}


plot(degree, cv.err5, type='b')

```

