---
title: "Project1 - Data 643"
author: "Ann Liu-Ferrara"
date: "June 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## the recommender system 

This system recommends data science books to readers.

 
## Data set selection and matrix 'user-item' 

```{r}
rating <-
  read.csv(
  'book_ratings.csv',
  na.strings = 'Did not read',
  stringsAsFactors = FALSE
  )[, -1]
  names(rating) <-
  c('name',
  'book1',
  'book2',
  'book3',
  'book4',
  'book5',
  'book6',
  'book7',
  'book8',
  'book9')
  
  # create a user-item matrix for each of train and test dataset
  set.seed(123)
  samp <- sample(nrow(rating), 0.7 * nrow(rating))
  train <- rating[samp, ]
  (rownames(train) <- train[, 1])
  (train <- as.matrix(train[, -1]))
  
  test <- rating[-samp, ]
  (rownames(test) <- test[, 1])
  (test <- as.matrix(test[, -1]))
  
  # Using training data, calculate the raw average (mean) rating for every user-item combination.
  (row.avg <- mean(train, na.rm = TRUE))
  
  # Calculate the RMSE for raw average for both your training data and your test data.
  (train.rmse <- sqrt(mean((train - row.avg)^2, na.rm=TRUE)))
  (test.rmse <- sqrt(mean((test - row.avg)^2, na.rm=TRUE)))
  
  # Using training data, calculate the bias for each user and each item.
  
  
  
```

# . From the raw average, and the appropriate user and item biases, calculate the baseline predictors for every user-item combination.
# . Calculate the RMSE for the baseline predictors for both your training data and your test data.
# . Summarize your results.
